
---
title: "Final Exam"
author: "DS 705"
output:
  word_document: default
  pdf_document: default
fontsize: 12pt
---

Unlike the Homework RMD files, this one doesn't contain all of the questions.  Rather we want you to read the questions on D2L and use this RMD file to record your R code that you used to answer each question.  This file must knit correctly.  After you submit the D2L quiz, then upload this RMD file and the knitted version of this file to the Final Exam dropbox on D2L.

## Questions 1-16

You do not need to submit any work with these questions.  Just answer them in D2L.

## Problem 1 - Questions 17-21

You're going to analyze the data in BirdPecks.rda and answer the questions in D2l.  Put all of your R in the chunk below:

```{r}
load("BirdPecks.rda")
boxplot(pecks~group , data = BirdPecks)

group1<-BirdPecks$pecks[BirdPecks$group=="1"]
group2<-BirdPecks$pecks[BirdPecks$group=="2"]

sd(group1)
sd(group2)

hist(group1,xlab="Number of Pecks",ylab="Frequency",main="Number of Pecks for Group 1")
hist(group2,xlab="Number of Pecks",ylab="Frequency",main="Number of Pecks for Group 2")

#test for normality
shapiro.test(group1)
shapiro.test(group2)
#both reject null hypothesis that distribution is normally distributed

#Problem 1 - Part D
wilcox.test(pecks~group,data=BirdPecks,conf.level=.9,conf.int=T)



```


---

## Problem 2 - Questions 22-26

Analyze StrengthSpeed.rda and put your R below:

```{r}
load("StrengthSpeed.rda")
boxplot(StrengthSpeed$before)
boxplot(StrengthSpeed$after)

differences<-StrengthSpeed$before-StrengthSpeed$after
boxplot(differences)
hist(differences)
#test numerically, not just visually
shapiro.test(differences)



```
$\mu_d$ is mean differences between groups. 
$$H_0: \mu_d = 0$$ 

$$H_a: \mu_d > 0$$


#Use R to conduct an appropriate hypothesis test to determine if strength training tends to increase speed in the target population. Use a 5% level of significance.

#and state the following: null and alternative hypotheses  test statistic  conclusion.
```{r}
t.test(StrengthSpeed$before,StrengthSpeed$after,alternative = "greater",paired=TRUE)

```
Here we reject the null hypothesis that there is no difference in means and accept the alternative that weight training does increase speed. (t = 9.6083, df = 34, p-value = 1.606e-11). 

---

## Problem 3 - Questions 27-36

Analyze GroupHLT scores and put your R here:

```{r}
win<-c(79, 76, 74, 70, 81, 85, 73, 78, 69, 72, 83, 89, 72, 79, 75)

loss<-c(78, 96, 85, 91, 77,103, 72, 93, 98, 86, 70, 110, 70, 91, 99)

#examining the data
boxplot(win);boxplot(loss)
hist(win);hist(loss)
shapiro.test(win);shapiro.test(loss )


mean(loss);mean(win)
t.test(win,loss)

t.test(loss)

library(boot)
bootMean<-function(x,i){
  mean(x[i])
}
#set seed to any integer value to preserve random number generation
set.seed(1);boot.object<-boot(loss,bootMean,R=5000)
boot.ci(boot.object,conf = .95)
```

---

## Problem 4 - Questions 37-41

Analyze the data in treadware.rda and put your R here:

```{r}
load("treadwear.rda")

```

---

## Problem 5 - Questions 42-59

Analyze the data in diamond.rda and include your R here:

```{r}
load("diamond.rda")
```

---

## Problem 6 - Questions 60- 66

Analyze the data in Shells.rda and include your R here:

```{r}
load("Shells.rda")
```

---

## Problem 7 - Questions 67-70

Analyze the "Primary News Source for Americans" data described in the problem statement.  Put your R below:

```{r}

```

---

## Problem 8 - Questions 71-75

Analyze the data in cheese.rda and put your R below:

```{r}
load("cheese.rda")
```

---

## Problem 9 - Questions 76-90

Analyze the data in careerbarrier.rda and put your R below:

```{r}
load("careerbarrier.rda")
```

---

## Problem 10 - Questions 91-98

Analyze the data on seat postion vs. nausea described in the problem.  Put your R below:

```{r}
```

---

## Question 99

Make sure both this RMD and the resulting knitted Word document are uploaded to the Dropbox "Final Exam R Code."
